Using SMT engine to generate Symbolic Automata\thanks{This work was partially funded by the Associated Team FM4CPS between INRIA and ECNU, Shanghai}

### 2018-01-19 2017-Xudong-SMT/FM-2018-extended.tex
reductions parties principales

\begin{table}\caption{\label{table:BIPalgebra}\leftline{Algebra
      Presentation: predefined and BIP Sorts and Operators}}
	\begin{tabular}{p{3cm}p{3cm}p{6cm}}
		\hline\specialrule{0em}{1pt}{1pt}
		Sort & Constructors & Other Operators
                \\\specialrule{0em}{1pt}{1pt}
		\hline\specialrule{0em}{3pt}{3pt}
		Bool    			&
                $\texttt{true},\ \texttt{false}$&
                $\land,\ \lor,\ \neg,\ \implies,\ =,\ \ne$
                \\\specialrule{0em}{1pt}{1pt} 
		Action 			&  $\texttt{Synchro},\ \texttt{FUN\_Action\_Bool}$ &
                \\\specialrule{0em}{1pt}{1pt}
		Int 				&
                ${0, \{i, -i\}_{i \in \texttt{Nat}}}$  &
                $- \texttt{(unary)},\ +,\ -
                \texttt{(binary)},\ \times,\ \div, \texttt{etc.}$
%                \\\specialrule{0em}{1pt}{1pt}
%                \textsl{for any sort} & & $=,\ \ne$
		\\\hline
	\end{tabular}
\end{table}

@@

\paragraph{Structural Semantic Rules:}
The semantics of pNet in term of open automata has been defined in
\cite{henrio:Forte2016}. For convenience we include these 
rules in appendix \ref{appendix:semRules}.

@@

Our algorithm starts with an open pNet, and builds its set of open
transitions (OTs). Its main loop is a classical residual algorithm: starting
from the initial global state, it pick a state in an unexplored set, and
computes all possible OTs, adding their target states in the
unexplored set, until this set is empty. As the potential set of
global states is finite (see \cite{henrio:Forte2016}), this terminates.

\begin{algorithm}[h]
  \caption{Open Automaton Generation}
  \label{alg1}
\begin{algorithmic}[1]
\Require The pNet node P.
\State Initialize sets $U$, $E$ for unexplored/explored global states, $L$ for result OTs;
\While{!isEmpty(U)}
	\State Chose $S$ in $U$;
	\State OT = MakeTransitions(P, S);
        \State Store $OT$ in $L$; 
	\For{each target global state $T$ from $OT \in L$}
		\IfThen{$(!U.contains(T))\ \&\&\ (!E.contains(T))$}{Add $T$ into the $U$;}
	\EndFor
\EndWhile
\State Prune the open transitions in $L$ using the SMT solver;
\State \Return $L$;

@@

The inside loop (\emph{MakeTransitions} method) applies recursively
the semantic rules following the structure of the pNet. 
For pLTS at leaves, rule Tr1 simply take the pLTS transitions of the
corresponding local state and build the predicate to generate the
open transition. 
On a pNet node to apply rule Tr2 we use two methods, combining and
matching, to generate the new predicate of the open transitions in a
hierarchical manner.

Last, at the root of the pNet, the predicate of each OT is translated
into SMTlib assertions, and checked for satisfiability. Unsatisfiable
OTs are pruned, and the final open automaton is computed as the set of
global states reachable by satisfiable OTs.

\begin{algorithm}
\caption{MakeTransitions()}
  \label{alg2}
\begin{algorithmic}[1]
\Require The pNet node P; The start global state S.
\State Initialize a list $l$, $L$ for sub-transitions/transitions.
	\For{each Subnet in P}
	\\ ~~~ $\setminus \setminus$ Recursively applying Tr1 or Tr2 on the sub-nodes.
                \State Store MakeTransitions(Subnet, S) in $l$;
	\EndFor
	\For{each $sv \in SV$}
		\State $comb = Combining(l)$;
                 \State $ot = Matching(sv, comb, hole, v)$;
                 \State Store $ot$ in $L$; 
	\EndFor
\State \Return $L$;
\end{algorithmic}  
\end{algorithm}

@@

\def\inactive{\{-\}}
\paragraph{Combining:}
The combining method enumerate all the possible behaviours of the
subnets, as all the possible combinations of their open transitions.
Assume that there is a collection of $n$ subnets $L = {s_1, s_2, ...,
  s_n}$, if we notate $ot_i$ to be the set of open transitions of the
subnet $s_i$ and ``$-$'' means the subnet is not involved.  
Then the combination $\texttt{COMB}$, a set of n-tuples, is the
cartesian product of sets :  
$\texttt{COMB} : (\inactive\cup ot_1) \times (\inactive\cup ot_2)\times ... \times (\inactive\cup ot_n) .$

@@

\paragraph{Matching:}
The next step, for each synchronisation vector, and each possible
combination of behaviour of the subnets as generated by the \emph
{Combining} method, consists in building the combined predicate,
and the corresponding open transition.
From a synchronization vector
$sv_k\in SV = (a'_i)^{i\in I} (b'_j)^{j\in J}\rightarrow v'$
and its guard $G_k$, a tuple of open transitions
$C_n\in \texttt{COMB} = {(v_i)}^{i\in I}$, the hole behaviors $Hole$,
and a fresh variable representing the result action $v$, we build the
predicate:

@@

Each of these possible failures is more precise than the previous
ones. But even when unification succeeds, it is still possible that
the resulting predicate would be unsatisfiable, because of some
incompatibility involving the guards. In our algorithm, we choose to
apply only the simplest filter inside the \emph{Matching} method. All
the rest will be checked later, using the satisfiability check in the
SMT engine.

@@

In a pLTS, there may be several incoming transitions 
that assign potentially different values to a state variable.
To handle such cases, the algorithm manages the assignments
within a list for each pLTS.
The list we used contains triples <$v$, $S$, $AssignRH$> 
where $v$ is the variable in pLTS, $S$ is its owner state and
$AssignRH$ is a list of expressions over other variables in the right
hand side of assignments of $v$.
% As a pragmatic extension to the formal definition, we also manage
% ``global variables'', defined in all states of the pLTS.
For a global state in the open automaton, getting the set of state
variables (which may be used in a transition) will simply be the union
of state variables of the individual pLTS states constituting this
global state (variables of different pLTS are distinct). We define:
\begin{itemize}
  \item $SVars(gs)$ the state variables of all states in the global
    state $gs$
  \item $Assigns(svar)$ the set of possible assignments of variable
      $svar$.
\end{itemize}

%There are some assignment on the variables in each of states of the pLTS (state variables), or just as global states of it (global variables). There is usually not only one assignment for a state variable, miss of them may drop the unsatisfiable results. In order to avoid such mistakes,

@@

Our matching/filtering strategy builds some transitions where
the predicates express incompatible constraints. Even if having an
unsatisfiable (symbolic) transition in the open automaton would not be incorrect,
we want to check them for satisfiability, and reduce the number of
transitions and states in the automaton.
In Fig.~\ref{schema:unsat-ot}, we display an example of an
unsatisfiable open transition 
from the result of our example. It shows the case where
the failure controller performs a ``$fail$'' action, while the timer
executes a ``$resume$''. The chosen synchronization
vector (SV0) does not match with these actions, as shown the predicate
fragment ``resume(true)=start($b_2$)''.

\begin{figure}[t]
\begin{mathpar}
    ot = \openrule{\{s1 \xrightarrow{fail(false)} s1, t1 \xrightarrow{resume(true)} t0\}, ~~\{\xrightarrow{hb}\},~~fail(false)=fail(b_1)\\
    \OTland resume(true)=start(b_2)\OTland hb=fail(b_0)\OTland v=\underline{fail}\OTland b_1=b_2\\
    \OTland (b_1\lor b_2) \Rightarrow b_0, ~~\{\}}
    	   {\ostate{s1,t1} \xrightarrow{v} \ostate{s1,t0}}\\
  \end{mathpar}
\caption{One of the unsatisfiable open transitions in FailureTimer}  \label{schema:unsat-ot}
\end{figure}

Checking satisfiability requires some symbolic computation
on the action expressions and the predicates, and this 
may depend on the specific theory of the action algebra. 
The ``Modulo Theory'' part of SMT solvers is important here, so that
the solver can use specific properties of each action algebra.

@@

As the last step of \ref{alg1}, we use the SMT solver Z3 to check the
satisfiability of each open transitions. More precisely, we translate
the OT predicate into a SMTlib assertion, in the context provided by
the possible state variable assignments in the source state. In this
section we describe the translation of the algebra presentation, of
assignments, and of the predicates.

@@

Before that work, we need to reconsider the result open transitions
from our algorithm. The ideal
result that we showed in the bottom part of
Fig.~\ref{schema:resultOA} is not the one generated by the algorithm,
because the original one contains many intermediate
variables, such as fresh local variables of synchronization
vectors, and fresh result variables of intermediate open transitions.
Such intermediate variables come from the way predicates are
generated, and depends on the particular structure of the pNet. But
they are not significant in the resulting behavior, and should be
eliminated before comparing such behavior with another pNet.
To eliminate the intermediate variables in the predicate, we want to define
the set of the significant variables that we won't eliminate: the
local variables of pLTS transitions, the hole behaviors and the result
action at the top level.
The result of the elimination will be simpler predicates, that can be
successfully compared (e.g. by Z3) within the bisimulation checking
algorithm.

@@

@@

\begin{definition}[Fresh variable]\label{fresh-variable}
The format of \emph{fresh variable} (renaming) is defined as:
``$\emph{prefix : tree\ index\ :\ counter}$''.
\begin{itemize}
   \item[$\bullet$] $prefix$ identifies the kind of the
     variable. current kinds are: sva (SV action), ra (result action), hb (hole behavior).
   \item[$\bullet$] $tree\ index$ is the index of the node containing the variable in the tree-like structure of the pNet.
   \item[$\bullet$] $counter$ is the current value of the corresponding counter.
\end{itemize}
\end{definition}

@@

For example, in the running example, the first possible behavior of
the hole, it doesn't have a name, only have the prefix ``hb''. The
hole is the second subnets of the root node so its ``tree index'' is
``12''. The fresh variable name is ``$\OTvar{:hb:12:1}$''.

@@

\section{Translation to SMTlib input language}
\label{appendix:mapsmt}
The pNet elements, as defined above, can be full translated into
SMT-LIB language, but there are a number of differences in the
structure of the models/languages, so the translation is not trivial.

@@

\QIN{Each time we submit each open transition to Z3 module, we translate
  its predicate into Z3 language format and send it for satisfiability
  checking. Every term of the predicate is declared as an
  \texttt{assert} in Z3. A constant action or a parameterized
  expression is easy to get from the internal list storing the objects
  while all the variables are not declared at the beginning. So we
  declare them before the submission of a predicate term with the API
  method conducting \texttt{declare-const}.}

@@

Let Pres = <Sorts, Constrs, Ops> be the extended presentation
and OT = <Leaves, Holes, Pred, Assign> an open transition.
\begin{itemize}
\item compute the environment $\EEnv=\EEnv_{pnet,OT}$ collecting
all variables used in OT.
\item check that all pLTS labels (action, guard, assignment) in
the transition, the OT predicate, and the OT assignments are
well-formed and well-typed.
\item for each start global state $gs$ of the open transition, using $SVar(gs)$
  get the state variables and their possible evaluations from all the state in $gs$.
  Then use Ass(x) as the set of evaluations of state variable x.
\item for each variable $v$ in $\EEnv$, construct:
  define-const = TrPredicate ($\EEnv ,v$).
%- turn the predicate into conjunctive normal form
\item for each conjunct $P_i$ in the predicate, construct:
  assert = TrPredicate $(P_i)$.
\end{itemize}

@@

\paragraph{Translating assignments into predicate terms}
\label{appendix:translating-assignments}
\TODO{INCORRECT, to be rewritten}
State-variable assignments are also treated as a part of the
predicates when checking the satisfiability. % Here we need conduct some
% translations,
For each assignement in a $\Post$ predicate, such as $\{s_0 \leftarrow 1\}$ in section
~\ref{section:op-semantics}, we translate it into an equation
$s_0=1$. For several assignments of the same variable in the same state,
we generate the disjunction of 
these equations. Correspondingly, we generate a conjunction for the
assignments from different variables.

%% \paragraph{Checking satisfiability}

%% % We check the satisfiability of open transitions on the top level of
%% % the pNet, treating these open transitions as the intermediate result
%% % of the algorithm. After we get the intermediate results of open
%% % transitions, we check them one by one.
%% Here we declare first all the variables in the predicates, with their
%% sorts.
%% An example of syntax is::\\
%% \centerline{$\OTvar{(declare-const\ |:ra:1:1|\ Action)}$}

%% \noindent Then each term of the
%% predicate is an assertion in Z3, as an example, in the former
%% unsatisfiable result in Fig~\ref{schema:unsat-ot}, the contradiction is occurred in terms: \\
%% \centerline{$\OTvar{(:ra:11:1=l)} \OTland \OTvar{(delta=:ra:11:1)}$}\\
%% from which we generate the following two assertions:\\
%% \centerline{$\OTvar{(assert\ (=\ |:ra:11:1|\ l))\ \ \ (assert\ (=\ delta\ |:ra:11:1|))}$}

%% \noindent The predicate submitted to the Z3 also contains the
%% assignments encodings.
%% In the case of $ot_2$, it
%% gives the assertion:\\
%% \centerline{$\OTvar{assert (or (= var\_s0 0) (= var\_s0 01)))}$}

@@

\begin{figure}[h]  
$\begin{array}{r@{}l}
    ot_1  = &\openrule{\{s0 \xrightarrow{\OTvar{fail(true)}} s1, t0 \xrightarrow{\OTvar{start(true)}} t1\}, ~~\{\xrightarrow{hb_1}\},~~\OTvar{fail(false)}=\OTvar{fail(}\OTvar{b}_1\OTvar{)}\\
    \OTland \OTvar{start(true)}=\OTvar{start(}\OTvar{b}_2\OTvar{)}\OTland hb_1=\OTvar{fail(}\OTvar{b}_0\OTvar{)}\\
    \OTland v_1=\OTvar{\underline{fail}}\OTland \OTvar{b}_1 = \OTvar{b}_2
    \OTland \OTvar{(}\OTvar{b}_1 \lor \OTvar{b}_2\OTvar{)} \Rightarrow \OTvar{b}_0, ~~\{t := Max\}}
    	   {\ostate{s0,t0} \xrightarrow{v_1} \ostate{s1,t1}}\\
\end{array}$

$\begin{array}{r@{}l}
    ot_2  = &\openrule{\{s0 \xrightarrow{\OTvar{resume(false)}} s0, t0 \xrightarrow{\OTvar{resume(false)}} t0\}, ~~\{\xrightarrow{hb_2}\},~~\OTvar{resume(false)}=\OTvar{resume(}\OTvar{b}_1\OTvar{)}\\
    \OTland \OTvar{resume(false)}=\OTvar{resume(}\OTvar{b}_2\OTvar{)}\OTland hb_2=\OTvar{resume(}\OTvar{b}_0\OTvar{)}\\
    \OTland v_2=\OTvar{\underline{resume}}\OTland \OTvar{b}_1 = \OTvar{b}_2
    \OTland \OTvar{(}\OTvar{b}_1 \lor \OTvar{b}_2\OTvar{)} \Rightarrow \OTvar{b}_0, ~~\{\}}
    	    {\ostate{s0,t0} \xrightarrow{v_2} \ostate{s0,t0}}\\
\end{array}$

>>>>

\begin{table}\caption{\label{table:BIPalgebra}\leftline{Algebra
      Presentation: predefined Sorts and Operators}}
	\begin{tabular}{p{3cm}p{3cm}p{6cm}}
		\hline\specialrule{0em}{1pt}{1pt}
		Sort & Constructors & Other Operators
                \\\specialrule{0em}{1pt}{1pt}
		\hline\specialrule{0em}{3pt}{3pt}
		Bool    			&
                $\texttt{true},\ \texttt{false}$&
                $\land,\ \lor,\ \neg,\ \implies,\ =,\ \ne$
                \\\specialrule{0em}{1pt}{1pt} 
                %               		Action 			&  $\texttt{Synchro},\ \texttt{FUN\_Action\_Bool}$ &
                Action 			&  $\texttt{Synchro},\ \texttt{FUN}$ &
                \\\specialrule{0em}{1pt}{1pt}
		Int 				&
                ${0, \{i, -i\}_{i \in \texttt{Nat}}}$  &
                $- \texttt{(unary)},\ +,\ -
                \texttt{(binary)},\ \times,\ \div, \texttt{etc.}$
%                \\\specialrule{0em}{1pt}{1pt}
%                \textsl{for any sort} & & $=,\ \ne$
		\\\hline\specialrule{0em}{1pt}{1pt}
\multicolumn{3}{l}{\sl Extension for the BIP use-case of Fig.~\ref{schema:ArchFailure:pNet}}
		\\\hline\specialrule{0em}{1pt}{1pt}
                Action & \multicolumn{2}{l} \texttt{FUN\_Action\_Bool}, \texttt{}fail, resume, timeout, reset, start, tick, ask 
		\\\hline
	\end{tabular}
\end{table}

@@

\paragraph{Structural Semantic Rules:}
The semantics of pNet in term of open automata has been defined in
\cite{henrio:Forte2016}, in the form of 2 structural rules, one for
pLTSs, one for pNet nodes.. For convenience we include these rules in
appendix \ref{appendix:semRules}.

@@

Our algorithm starts with an open pNet, and builds its set of open
transitions. Its main loop is a classical residual algorithm: starting
from the initial global state, it picks a state in an unexplored set, and
computes all possible OTs, adding their target states in the
unexplored set, until this set is empty. As the potential set of
global states is finite (see \cite{henrio:Forte2016}), this terminates.

\begin{algorithm}[h]
  \caption{Open Automaton Generation}
  \label{alg1}
\begin{algorithmic}[1]
\Require The pNet node P.
\State Initialize sets $U=InitSate(P)$, $E=\emptyset$ for unexplored/explored global states, $L$ for result OTs;
\While{!isEmpty(U)}
	\State Chose $S$ in $U$;
	\State OT = MakeTransitions(P, S);
        \State Store $OT$ in $L$; 
	\For{each target global state $T$ from $OT \in L$}
		\IfThen{$(!U.contains(T))\ \&\&\ (!E.contains(T))$}{Add $T$ into the $U$;}
	\EndFor
\EndWhile
\State Prune the open transitions in $L$ using the SMT solver;
\State \Return $OA=Reachable(InitState(P),L)$;

@@

The inside loop (\emph{MakeTransitions} method) applies recursively
the semantic rules following the structure of the pNet. 
For pLTS at leaves, we simply take the pLTS transitions of the
corresponding local state and build the predicate to generate the
open transition. 
On a pNet node we use two methods, combining and
matching, to generate the predicate of the open transitions in a
hierarchical manner.

Then, at the root of the pNet, the predicate of each OT is translated
into SMTlib assertions, and checked for satisfiability. Unsatisfiable
OTs are pruned, and the final open automaton is computed as the set of
global states reachable by satisfiable OTs.

\begin{algorithm}
\caption{MakeTransitions()}
  \label{alg2}
\begin{algorithmic}[1]
\Require a pNet node Pn; a global state S.
\State Initialize a list $l$, $L$ for sub-transitions/transitions.
	\For{each Subnet in Pn}
	~~~ $\setminus \setminus$ Recursively applying Tr1 or Tr2 on the sub-nodes.
                \State Store MakeTransitions(Subnet, S) in $l$;
	\EndFor
	\For{each $sv \in SV$}
		\State $\overline{comb} = Combining(l)$;
                 \State $ot = Matching(sv, \overline{comb}, hole, v)$;
                 \State Store $ot$ in $L$; ~~~ $\setminus \setminus$ if Filter(ot) succeeds
	\EndFor
\State \Return $L$;
\end{algorithmic}  
\end{algorithm}

@@

\def\inactive{\{-\}}
\paragraph{Combining:}
The combining method enumerate all the possible behaviours of the
subnets, as all the possible combinations of their open transitions.
Assume that there is a collection of $n$ subnets
$L = {s_1, s_2, ...,  s_n}$; we denote $\overline{ot}_i$ the set of open
transitions of the 
subnet $s_i$ and ``$-$'' means the subnet is not involved;
then the combination $\overline{comb}$, a set of n-tuples, is the
cartesian product of sets :  
$\overline{comb} : (\inactive\cup \overline{ot}_1) \times (\inactive\cup \overline{ot}_2)\times ... \times (\inactive\cup \overline{ot}_n).$

@@

\paragraph{Matching:}
The next step, for each synchronisation vector, and each possible
combination of behaviour of the subnets as generated by the \emph
{Combining} method, consists in building the combined predicate,
and the corresponding open transition.
From a synchronization vector
$sv_k\in SV = (a'_i)^{i\in I} (b'_j)^{j\in J}\rightarrow v'$
and its guard $G_k$, a tuple of open transitions
$C_n\in \overline{comb} = {(v_i)}^{i\in I}$, the hole behaviors $Hole$,
and a fresh variable representing the result action $v$, we build the
predicate:

@@

Each of these possible filters is more precise than the previous
ones. But even when unification succeeds, it is still possible that
the resulting predicate would be unsatisfiable, because of some
incompatibility involving the guards. In our algorithm, we choose to
apply only the simplest filter inside the \emph{Matching} method. All
the rest will be checked later, using the satisfiability check in the
SMT engine.

@@

In a pLTS, there may be several incoming transitions 
that assign potentially different values to a state variable.
To handle such cases, the algorithm manages for each pLTS state, as a
list of expressions collected from the assignments of each state
variable $v$.
%% The list we used contains triples <$v$, $S$, $AssignRH$> 
%% where $v$ is the variable in pLTS, $S$ is its owner state and
%% $AssignRH$ is a list of expressions over other variables in the right
%% hand side of assignments of $v$.
% As a pragmatic extension to the formal definition, we also manage
% ``global variables'', defined in all states of the pLTS.
For a global state in the open automaton, getting the set of state
variables (which may be used in a transition) will simply be the union
of state variables of the individual pLTS states constituting this
global state (variables of different pLTS are distinct).
%There are some assignment on the variables in each of states of the pLTS (state variables), or just as global states of it (global variables). There is usually not only one assignment for a state variable, miss of them may drop the unsatisfiable results. In order to avoid such mistakes,

@@

Our matching/filtering strategy builds some transitions where
the predicates express incompatible constraints. Even if having an
unsatisfiable (symbolic) transition in the open automaton would not be incorrect,
we want to check them for satisfiability, and reduce the number of
transitions and states in the automaton.
In Fig.~\ref{schema:unsat-ot}, we display an example of an
unsatisfiable open transition 
from the result of our example. It shows the case where
the failure controller performs a ``$fail$'' action, while the timer
executes a ``$resume$''. The chosen synchronization
vector (SV0 from Fig.~\ref{schema:ArchFailure:pNet}) does not match
with these actions, as shown by the predicate
fragment ``resume(true)=start($b_2$)''.

\begin{figure}[t]
\begin{mathpar}
    ot = \openrule{\{s1 \xrightarrow{fail(false)} s1, t1 \xrightarrow{resume(true)} t0\}, ~~\{\xrightarrow{hb}\},~~fail(false)=fail(b_1)\\
    \OTland resume(true)=start(b_2)\OTland hb=fail(b_0)\OTland v=\underline{fail}\OTland b_1=b_2\\
    \OTland (b_1\lor b_2) \Rightarrow b_0, ~~\{\}}
    	   {\ostate{s1,t1} \xrightarrow{v} \ostate{s1,t0}}\\
  \end{mathpar}
\caption{One of the unsatisfiable open transitions in the Failure Monitor pNet}  \label{schema:unsat-ot}
\end{figure}

Checking satisfiability requires some symbolic computation
on the action expressions and the predicates, and this 
may depend on the specific theory of the action algebra datatypes. 
The ``Modulo Theory'' part of SMT solvers is important here, so that
the solver can use specific properties of each action algebra.

@@

We check the satisfiability of each open transitions
using the SMT solver Z3. More precisely, we translate
the OT predicate into a SMTlib assertion, in the context provided by
the possible state variable assignments in the source state. In this
section we describe the translation of the algebra presentation, of
assignments, and of the predicates.

@@

@@

In rule TR2, the generated predicate is composed of the conjunction of
the predicates of the subnets' OTs, with the additional part encoding
the application of the chosen synchronisation vector.
In \cite{henrio:Forte2016} this last part is defined as:
\[\MkPred(SV_k, a_i^{i\in I}, b_j^{j\in J}, v)\Leftrightarrow
\begin{array}{l}
\exists {(a'_i)}^{i\in I},
{(b'_j)}^{j\in J},v'.\, SV_k={(a'_i)}^{i\in I}, {(b'_j)}^{j\in J}\rightarrow v'
\\~~\land
\forall i\in I.\, a_i=a'_i\land \forall j \in J.\, b_j=b'_j \land v=v'
\end{array}\]

This definition is not constructive, but the existentialy quantified
variable sets here correspond the subsets of subnets and holes that
are effectively involved in vector $SV_k$.
Within our algorithm, these subsets have been computed by $Combining$ and
passed as arguments to the $Matching$ method, so
we can construct this predicate fragment as:

\[\Pred(SV, \overline{\pNet},\overline{S}, v) = 
\begin{array}{l}
\forall {(a'_i)}^{i\in I},
{(b'_j)}^{j\in J},v'.
\\
v_m=a'_i\land \, b_j=b'_j \land v=v' \land G_k
\end{array}\]
\TODO{Eric: to be checked}

@@

\begin{definition}[Fresh variable]\label{fresh-variable}
The format of \emph{fresh variable} (renaming) is defined as:
``$\emph{prefix : tree\ index\ :\ counter}$''.
\begin{itemize}
   \item[$\bullet$] $prefix$ identifies the kind of the
     variable. Current kinds are: sva\_<SV> (action variable in vector
     $SV$), ra (result action), hb\_<B> (behavior of hole $B$).
   \item[$\bullet$] $tree\ index$ is the index of the node containing
     the variable in the tree-like structure of the pNet. 
   \item[$\bullet$] $counter$ is the current value of the corresponding counter.
\end{itemize}
\end{definition}

@@

For example, in the running example, in the raw output listing in
Appendix~ref{}, in the first OT, you find variables: 
\begin{itemize}
   \item $\OTvar{:hb\_B:13:1}$ that is the behaviour of hole ``B''
   \item $\OTvar{b1:sva\_SV0:1:1}$ variable ``b1'' from vector ``SV0''
   \item $\OTvar{:ra:1:1}$ result action of the current OT
\end{itemize}

@@

\section{Management of state variables}
\label{appendix:translating-assignments}
In a pLTS, there may be several incoming transitions 
that assign potentially different values to a state variable.
To handle such cases, the algorithm manages for each pLTS state, as a
list of expressions collected from the assignments of each state
variable $v$ (as registered in the incoming OTs).
%% The list we used contains triples <$v$, $S$, $AssignRH$> 
%% where $v$ is the variable in pLTS, $S$ is its owner state and
%% $AssignRH$ is a list of expressions over other variables in the right
%% hand side of assignments of $v$.
% As a pragmatic extension to the formal definition, we also manage
% ``global variables'', defined in all states of the pLTS.
For a global state in the open automaton, getting the set of state
variables (which may be used in a transition) will simply be the union
of state variables of the individual pLTS states constituting this
global state, as variables of different pLTS are distinct.

 We define:
\begin{itemize}
  \item $SVars(gs)$ the state variables of all states in the global
    state $gs$
  \item $Assigns(svar)$ the set of possible assignments of variable
      $svar$.
\end{itemize}
These will be used when translating the predicates into SMTlib
language.

\section{Translation to SMTlib input language}
\label{appendix:mapsmt}
The pNet elements, as defined above, can be full translated into
SMT-LIB language (\ref{SMTlibv2.5}), but there are a number of differences in the
structure of the models/languages, so the translation is not trivial.

@@

Each time we submit each open transition to Z3 module, we translate
  its predicate into the SMTlib language  and send it for satisfiability
  checking. Every term of the predicate is declared as an
  \texttt{assert} in Z3. A constant action or a parameterized
  expression is easy to get from the internal list storing the objects
  while all the variables are not declared at the beginning. So we
  declare them before the submission of a predicate term with the API
  method conducting \texttt{declare-const}.

@@

Let Pres = <Sorts, Constrs, Ops> be the extended presentation
and OT = <Leaves, Holes, Pred, Assign> an open transition.
\begin{itemize}
\item compute the environment $\EEnv=\EEnv_{pnet,OT}$ collecting
all variables used in OT.
\item check that all pLTS labels (action, guard, assignment) in
the transition, the OT predicate, and the OT assignments are
well-formed and well-typed.
\item for each start global state $gs$ of the open transition, using $SVar(gs)$
  get the state variables and their possible evaluations from all the state in $gs$.
  Then use Assigns(x) as the set of evaluations of state variable x.
\item for each variable $v$ in $\EEnv$, construct:
  define-const = TrPredicate ($\EEnv ,v$).
%- turn the predicate into conjunctive normal form
\item for each conjunct $P_i$ in the predicate, construct:
  assert = TrPredicate $(P_i)$.
\end{itemize}

@@

\subsection{properties}
The rules in the sections above are meant to guarantee that no runtime
errors would occur when running Z3 from our algorithm. More precisely,
we ensure that:
\begin{itemize}
\item all datatypes, functions, variable used are declared,
\item datatypes declared are well-founded,
\item all expressions in declarations and assertions are well-typed.
\end{itemize}

\newpage
\begin{figure}[h]
$\begin{array}{r@{}l}
    ot_1  = &\openrule{\{s0 \xrightarrow{\OTvar{fail(true)}} s1, t0 \xrightarrow{\OTvar{start(true)}} t1\}, ~~\{\xrightarrow{hb_1}\},~~\OTvar{fail(false)}=\OTvar{fail(}\OTvar{b}_1\OTvar{)}\\
    \OTland \OTvar{start(true)}=\OTvar{start(}\OTvar{b}_2\OTvar{)}\OTland hb_1=\OTvar{fail(}\OTvar{b}_0\OTvar{)}\\
    \OTland v_1=\OTvar{\underline{fail}}\OTland \OTvar{b}_1 = \OTvar{b}_2
    \OTland \OTvar{(}\OTvar{b}_1 \lor \OTvar{b}_2\OTvar{)} \Rightarrow \OTvar{b}_0, ~~\{t := Max\}}
    	   {\ostate{s0,t0} \xrightarrow{v_1} \ostate{s1,t1}}\\
\end{array}$
  \smallskip

$\begin{array}{r@{}l}
    ot_2  = &\openrule{\{s0 \xrightarrow{\OTvar{resume(false)}} s0, t0 \xrightarrow{\OTvar{resume(false)}} t0\}, ~~\{\xrightarrow{hb_2}\},~~\OTvar{resume(false)}=\OTvar{resume(}\OTvar{b}_1\OTvar{)}\\
    \OTland \OTvar{resume(false)}=\OTvar{resume(}\OTvar{b}_2\OTvar{)}\OTland hb_2=\OTvar{resume(}\OTvar{b}_0\OTvar{)}\\
    \OTland v_2=\OTvar{\underline{resume}}\OTland \OTvar{b}_1 = \OTvar{b}_2
    \OTland \OTvar{(}\OTvar{b}_1 \lor \OTvar{b}_2\OTvar{)} \Rightarrow \OTvar{b}_0, ~~\{\}}
    	    {\ostate{s0,t0} \xrightarrow{v_2} \ostate{s0,t0}}\\
  \end{array}$
\smallskip

$\begin{array}{r@{}l}
    ot_3  = &\openrule{\{\}, ~~\{\xrightarrow{hb_3}\},~~hb_3=\OTvar{finish}\OTland v_3=\OTvar{finish}, ~~\{\}}
    	    {\ostate{s0,t0} \xrightarrow{v_3} \ostate{s0,t0}}\\
\end{array}$
\smallskip

$\begin{array}{r@{}l}
    ot_4  = &\openrule{\{s1 \xrightarrow{\OTvar{fail(false)}} s1, t1 \xrightarrow{\OTvar{start(false)}} t1\}, ~~\{\xrightarrow{hb_4}\},~~\OTvar{fail(false)}=\OTvar{fail(}\OTvar{b}_1\OTvar{)}\\
    \OTland \OTvar{start(false)}=\OTvar{start(}\OTvar{b}_2\OTvar{)}\OTland hb_4=\OTvar{fail(}\OTvar{b}_0\OTvar{)}\\
    \OTland v_4=\OTvar{\underline{fail}}\OTland \OTvar{b}_1 = \OTvar{b}_2
    \OTland \OTvar{(}\OTvar{b}_1 \lor \OTvar{b}_2\OTvar{)} \Rightarrow \OTvar{b}_0, ~~\{\}}
    	   {\ostate{s1,t1} \xrightarrow{v_4} \ostate{s1,t1}}\\
\end{array}$
\smallskip

$\begin{array}{r@{}l}
    ot_5  = &\openrule{\{s1 \xrightarrow{\OTvar{resume(true)}} s0, t1 \xrightarrow{\OTvar{resume(true)}} t0\}, ~~\{\xrightarrow{hb_5}\},~~\OTvar{resume(true)}=\OTvar{resume(}\OTvar{b}_1\OTvar{)}\\
    \OTland \OTvar{resume(true)}=\OTvar{resume(}\OTvar{b}_2\OTvar{)}\OTland hb_2=\OTvar{resume(}\OTvar{b}_0\OTvar{)}\\
    \OTland v_5=\OTvar{\underline{resume}}\OTland \OTvar{b}_1 = \OTvar{b}_2
    \OTland \OTvar{(}\OTvar{b}_1 \lor \OTvar{b}_2\OTvar{)} \Rightarrow \OTvar{b}_0, ~~\{\}}
    	    {\ostate{s1,t1} \xrightarrow{v_5} \ostate{s0,t0}}\\
\end{array}$
\smallskip

$\begin{array}{r@{}l}
    ot_6  = &\openrule{\{t1 \xrightarrow{\OTvar{tick}} t1\}, ~~\{\},~~t\neq 0\OTland v_6=\OTvar{tick}, ~~\{t := t-1 \}}
    	   {\ostate{s1,t1} \xrightarrow{v_6} \ostate{s1,t0}}\\
\end{array}$
\smallskip

$\begin{array}{r@{}l}
    ot_7  = &\openrule{\{\}, ~~\{\xrightarrow{hb_7}\},~~hb_7=\OTvar{finish}\OTland v_7=\OTvar{finish}, ~~\{\}}
    	    {\ostate{s1,t1} \xrightarrow{v_7} \ostate{s1,t1}}\\
\end{array}$
\smallskip

$\begin{array}{r@{}l}
    ot_8  = &\openrule{\{s1 \xrightarrow{\OTvar{timeout}} s2, t1 \xrightarrow{\OTvar{timeout}} t0\}, ~~\{\},~~t=0\OTland v_8=ask, ~~\{\}}
    	   {\ostate{s1,t1} \xrightarrow{v_8} \ostate{s2,t0}}\\
\end{array}$
\smallskip

$\begin{array}{r@{}l}
    ot_9  = &\openrule{\{s2 \xrightarrow{\OTvar{resume(false)}} s2, t0 \xrightarrow{\OTvar{resume(false)}} t0\}, ~~\{\xrightarrow{hb_9}\},~~\OTvar{resume(false)}=\OTvar{resume(}\OTvar{b}_1\OTvar{)}\\
    \OTland \OTvar{resume(false)}=\OTvar{resume(}\OTvar{b}_2\OTvar{)}\OTland hb_9=\OTvar{resume(}\OTvar{b}_0\OTvar{)}\\
    \OTland v_9=\OTvar{\underline{resume}}\OTland \OTvar{b}_1 = \OTvar{b}_2
    \OTland \OTvar{(}\OTvar{b}_1 \lor \OTvar{b}_2\OTvar{)} \Rightarrow \OTvar{b}_0, ~~\{\}}
    	    {\ostate{s2,t0} \xrightarrow{v_9} \ostate{s2,t0}}\\
\end{array}$
\smallskip

$\begin{array}{r@{}l}
    ot_{10}  = &\openrule{\{s2 \xrightarrow{\OTvar{reset}} s0\}, ~~\{\},~~v_{10}=\OTvar{reset}, ~~\{\}}
    	   {\ostate{s2,t0} \xrightarrow{v_{10}} \ostate{s0,t0}}\\
\end{array}$
\smallskip

\newpage
\subsection{Scaling up: several subsystems on the same bus}

@@
